### 2022-02-03 plos_latex_template.tex
co-author edits incorporated by @AndreaSanchezTapia

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5,6},
Heather Turner\textsuperscript{7}, %\ddag},
Dorothea Hug Peter\textsuperscript{8}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{9},
Matt Bannert\textsuperscript{10},
Batool Almazrouq\textsuperscript{11,12},
Liz Hare\textsuperscript{13},
Laura Ación\textsuperscript{5,14},
Juan Pablo Narváez-Gómez\textsuperscript{15},
Marcela Alfaro Córdoba\textsuperscript{16},
Federico Marini\textsuperscript{17},
Rita Giordano\textsuperscript{18},
Adithi Upadhya\textsuperscript{19},
Joselyn Chávez\textsuperscript{20},
Anicet Ebou\textsuperscript{21},
Silvia Canelón\textsuperscript{22},
Janani Ravi\textsuperscript{23}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{7} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{8} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{9} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{10} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{11} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
\\
\textbf{12} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{13} Affiliation Dept/Program/Center, Institution Name, City, State, Country
\\
\textbf{14} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{15} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
\\
\textbf{16} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{17} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{18} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{19} ILK Labs/Bengaluru, Karnataka 560046, India
\\
\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
\\
\textbf{21} Bioinformatic team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{22} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
\\
\textbf{23} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

% Conferences are great, but only for some
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
However, opportunities for participating in conferences are not equal for all. 
Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income countries, prestigious institutions, native English speakers with no physical disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
We had the opportunity to organize a virtual and global statistical computing conference, useR! 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.

% Roadmap of the paper
The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
They are organized in three sections: three pillars, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
\textbf{The pillars} comprise fundamental rules with key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
\textbf{Rule 8} provides tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
a \textbf{continuity Rule} emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion.

@@

\begin{figure}[!h]
\centering
%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file. ask how to do this this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
\pdftooltip{
\includegraphics[width=\textwidth]{figs/rules.png}}{
%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: pillar rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10). The overall diagram depicts an arrow from left to right which contains all the rules. From left to right, "the pillars" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team". Then, the label "Design Rules" is surrounded by six blue bubbles with Rules "4 Counteract bias", "5 Online", "6 Accessibility", "7 Language", "8 Communication",  "9 Financial". To the right, and forming the head of the arrow, the Tenth Rule is under the label "Continuity" and consists of a blue bubble "10 Assess and pass along".
}
\caption{Schematic diagram of the rules organized in three groups: the pillars (Rules 1, 2, and 3), design rules (Rules 4 to 9), and continuity (Rule 10).}
\label{fig:diagram}
\end{figure}

@@

Diversity encompasses multiple dimensions: age, physical ability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. 
Nonetheless, we live in a world with implicit hierarchies along these axes. 
Some statuses (e.g. being cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built. 
People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways.
As a result, some groups of people have been systematically excluded from or only partially included in academia and scientific and professional circles.
Recognizing this systematic exclusion is paramount to inclusion.
Assess how it translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
Define measurable indicators for these goals (e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions) that will help you be accountable along the way.
The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.

@@

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects you committed to in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of preferred pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and  great advice on `low-hanging fruits').
Open the schedule for events organized by thematic groups and communities (e.g. task force meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.

@@

After experiencing virtual conferences mostly due to the COVID-19 pandemic, there have been multiple calls to retain an online component in conferences, either as completely virtual events or using hybrid formats (combining in-person and online components) \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
Virtual conferences remove barriers for inclusion like costs of participation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}. 
Make sure that you are not creating additional barriers when choosing conference platforms, schedule, and activities. 
Check every platform's screen reader accessibility (see \textbf{Rule 6}), bandwidth requirements (a high-speed internet prerequisite could exclude attendees from many regions), and geopolitical restrictions (some services are not available in all countries).
Be mindful of the different time zones and time constraints of the potential attendees and create flexible schedules.
Avoid creating a strictly synchronous live program, ask for pre-recorded talks, record sessions, and make them available during the conference.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. Including people with disabilities in the organizing team from the start can have a large impact (see \textbf{Rule 3}), since planning for accessibility requires time, experience, early decision-making, and inaccessible features are difficult to correct at later stages \cite{irishIncreasingParticipationUsing2020}.

If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, have signs in Braille, and a sound system compatible with hearing devices and language interpretation, to name a few. 
Invisible disabilities  (e.g. dyslexia, anxiety, ADHD, autism) should be accounted for proactively, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.

@@

Encourage the preparation of accessible slides and presentations by providing accessibility guidelines and be available for any questions that presenters and attendees may have; see \cite{sanchez-tapia_user_2021} for an example of accessibility guidelines and \cite{chavez_preparing_2021, sanchez-tapia_making_2021, joo_how_2021} for further accessibility recommendations.
Social events and networking should also have accessibility features like captioning or sign language interpretation, and include activities that do not restrict participation based on body type or ability.

@@

The communication strategy of your conference should aim to reach broader audiences and express the commitment with diversity and inclusion and the welcoming spirit of the conference.
Actively reach out and promote the conference to communities that have been systematically excluded. 
Promote the abstract submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
The members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists).%#promotion2

@@

Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups—in all communications \cite{hallDesigningDiversityInclusion2019}. 
Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group and the terms that should be avoided.
Do not expect minoritized people to teach you and accept feedback without being offended.
Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
Resources permitting, enable a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from marginalized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
Simplifying the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the virtual component. 
Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees, give gift cards). Depending on the country, services like OpenCollective (\url{https://opencollective.com}) can be very useful to do these transfers.

@@

Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, or personal issues affecting your organizing team. 
Try to account for these factors as best as you can to set a realistic yet bold vision and goals of diversity and inclusion for the conference.
At the end of the day, even if you do not achieve them at 100\%, striving for them will make you stretch what feels possible and help you make changes you would have otherwise not even attempted to make. 
And most importantly, you would have contributed to a great conference experience for the attendees. 
They will notice the welcoming and inclusive spirit of the conference reflected in the pillar rules and the effort put in every aspect of the conference design. 
Inclusive conferences can have lasting impacts in career paths and create a sense of belonging in the community. 
That is more than enough motivation to make the effort worth it.

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 (\url{https://user2021.r-project.org/about/global-team}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process. Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten simple rules to host an inclusive conference} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Rocío Joo\textsuperscript{1*\Yinyang}, 
Andrea Sánchez-Tapia\textsuperscript{2\Yinyang},
Sara Mortara\textsuperscript{3},
Yanina Bellini Saibene\textsuperscript{4,5,6},
Heather Turner\textsuperscript{7}, %\ddag},
Dorothea Hug Peter\textsuperscript{8}, %\ddag},
Natalia Soledad Morandeira\textsuperscript{9},
Matt Bannert\textsuperscript{10},
Batool Almazrouq\textsuperscript{11,12},
Liz Hare\textsuperscript{13},
Laura Ación\textsuperscript{5,14},
Juan Pablo Narváez-Gómez\textsuperscript{15},
Marcela Alfaro Córdoba\textsuperscript{16},
Federico Marini\textsuperscript{17},
Rita Giordano\textsuperscript{18},
Adithi Upadhya\textsuperscript{19},
Joselyn Chávez\textsuperscript{20},
Anicet Ebou\textsuperscript{21},
Silvia Canelón\textsuperscript{22},
Janani Ravi\textsuperscript{23}
%,3\textcurrency}
\\
\bigskip
\textbf{1} Global Fishing Watch, Washington, DC 20036, USA
\\
\textbf{2} Instituto de Pesquisas Jardim Botânico do Rio de Janeiro, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{3} International Institute for Sustainability, Rio de Janeiro, Rio de Janeiro, Brazil
\\
\textbf{4} Instituto Nacional de Tecnología Agropecuaria, Ruta Nac. N° 5, Km. 580, Anguil, La Pampa, Argentina
\\
\textbf{5} MetaDocencia, Argentina
\\
\textbf{6} Universidad Nacional Guillermo Brown, Buenos Aires, Argentina
\\
\textbf{7} Department of Statistics, University of Warwick, Coventry, UK
\\
\textbf{8} Swiss Federal Institute for Forest, Snow and Landscape Research WSL, Birmensdorf, Switzerland
\\
\textbf{9} Instituto de Investigación e Ingeniería Ambiental, Universidad Nacional de San Martín - Consejo Nacional de Investigaciones Científicas y Técnicas, General San Martín, Buenos Aires, Argentina
\\
\textbf{10} KOF Swiss Economic Institute, ETH Zurich, Zurich, Switzerland
\\
\textbf{11} Institute of Systems, Molecular and Integrative Biology, University of Liverpool, Liverpool, L69 7ZB, UK
\\
\textbf{12} King Abdullah International Medical Research Center, Riyadh, Saudi Arabia
\\
\textbf{13} Dog Genetics LLC, Astoria, NY, USA
\\
\textbf{14} Instituto de Cálculo, FCEyN, UBA-CONICET, Buenos Aires, Argentina
\\
\textbf{15} Departamento de Botânica, Instituto de Biociências, Universidade de São Paulo, São Paulo, São Paulo, Brasil
\\
\textbf{16} Department of Statistics, University of California, Santa Cruz, CA, USA
\\
\textbf{17} Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), University Medical Center Mainz, Mainz, Germany
\\
\textbf{18} Royal Society of Chemistry, Thomas Graham House (290), Science Park, Milton Road, Cambridge, UK
\\
\textbf{19} ILK Labs/Bengaluru, Karnataka 560046, India
\\
\textbf{20} Instituto de Biotecnología, UNAM, Cuernavaca, Morelos, México
\\
\textbf{21} Bioinformatics team, Département de Formation et de Recherches en Agriculture et Ressources Animales, Institut National Polytechnique Félix Houphouët-Boigny, BP 1313 Yamoussoukro, Côte d’Ivoire.
\\
\textbf{22} Department of Biostatistics, Epidemiology \& Informatics, University of Pennsylvania, Philadelphia, PA, USA.
\\
\textbf{23} Departments of Pathobiology and Diagnostic Investigation, Microbiology and Molecular Genetics, Michigan State University, East Lansing, MI 48824, USA
\\
\bigskip

@@

% Conferences are great, but only for some
Conferences are spaces to meet and reconnect with members from a specific community, learn about advances in the field, and share our recent contributions.
A good conference experience can make a difference in the professional development of the participants and create long-lasting collaborations and opportunities. 
However, opportunities for participating in conferences are not equal for all. 
Many academic and tech conferences have become spaces that reproduce systemic inequalities, by failing to eliminate the barriers for participation and giving more opportunities to the most privileged individuals (typically white people from high-income countries, prestigious institutions, native English speakers with no disabilities, for example) \cite{arendDisparityConferenceRegistration2019, biggsAcademicConferenceChilly2018, depickerRethinkingInclusionDisability2020a, irishIncreasingParticipationUsing2020}.
The authors of this article have experienced these inequalities in conferences, from different perspectives and levels of inclusion/exclusion and privilege.
We had the opportunity to organize a virtual and global statistical computing conference, useR! 2021 \cite{sanchez-tapia_user_2021-2}, for users and developers of the R programming language \cite{r_core_team_2021}.
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.

% Roadmap of the paper
The rules result from our lessons learned before, during, and after the organization of the useR! conference. 
They are organized in three sections: three foundational rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundational rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and other spaces). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to encourage and ensure full participation of non-native English speakers. 
\textbf{Rule 8} provides tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion.

@@

\begin{figure}[!h]
\centering
%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file. ask how to do this this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
\pdftooltip{
\includegraphics[width=\textwidth]{figs/rules.png}}{
%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundational rules (rules 1-3), design rules (rules 4 to 9), and a continuity rule (rule 10). The overall diagram depicts an arrow from left to right which contains all the rules. From left to right, "the foundations" label is surrounded by blue bubbles which show Rule "1 Vision", Rule "2 Safe environment", and Rule "3 Organizing Team". Then, the label "Design Rules" is surrounded by six blue bubbles with Rules "4 Counteract bias", "5 Online", "6 Accessibility", "7 Language", "8 Communication",  "9 Financial". To the right, and forming the head of the arrow, the Tenth Rule is under the label "Continuity" and consists of a blue bubble "10 Assess and pass along".
}
\caption{Schematic diagram of the rules organized in three groups: the foundations (Rules 1, 2, and 3), conference design (Rules 4 to 9), and continuity (Rule 10).}
\label{fig:diagram}
\end{figure}

@@

Diversity encompasses multiple dimensions: age, disability, career stage, gender, gender identity, geographic origin, language, neurodiversity, race, religion, sexual orientation, and socioeconomic background, to name a few.
Human diversity should be celebrated and respected in every way. 
Nonetheless, we live in a world with implicit hierarchies along these axes. 
Some statuses (e.g. being cisgender, white, male, from North America or Western Europe) hold the privilege of being the defaults around which all systems—including conferences—are consciously and unconsciously built. 
People outside the dominant groups suffer different forms of oppression, like sexism, racism, ableism, homophobia, and transphobia, and sometimes these forms of oppression happen simultaneously and in complex ways \cite{crenshawDemarginalizingIntersectionRace1989}.
As a result, some groups of people have been systematically excluded from or only partially included in academia and scientific and professional circles.
Recognizing this systematic exclusion is paramount to inclusion.
Assess how it translates to your particular field and scientific or professional community and use this assessment to set the vision and goals of diversity and inclusion for your conference.
Define measurable indicators for these goals (e.g. a gender distribution of your speakers that is representative of the general population, the participation of people from diverse races and ethnicities among the organizers, speakers, and attendees, or participation from key geographic regions) that will help you be accountable along the way.
The vision can be expressed in a diversity statement to be used as a reference for decision making in all areas of the conference organization.

@@

% Inclusion for everyone
% Explain why we need this
Real inclusion can only be achieved in an environment where everybody feels safe, respected, and invited to take part of the conference activities and community. Be proactive and design a welcoming conference experience that takes into account the well-being of all the attendees. This means respecting all the aspects you committed to in your diversity vision (see \textbf{Rule 1}). For example, promote the use and respect of pronouns and commit to respect all genders and sexual orientations. If the conference has an in-person component, pre-assign quiet, private spaces for religious practices, lactation, and child care. Include menu options that account for diverse dietary and religious requirements.
Some of these actions are very easy to implement and have a significant impact on making everyone feel truly included (see \cite{numfocus_discover_2021} for more examples and  great advice on `low-hanging fruits').
Open the schedule for events organized by thematic groups and communities (e.g. task force meetings, LGBTQIA+-friendly spaces, Black in STEM), and sessions aimed to welcome first-timers to the conference.

@@

After experiencing virtual conferences mostly due to the COVID-19 pandemic, there have been multiple calls to retain an online component in conferences, either as completely virtual events or using hybrid formats (combining in-person and online components) \cite{jooKeepOnlineOption2021, woolstonLearningLoveVirtual2020, ninerBetterWhomLeveling2021, roosOnlineConferencesNew2020, levitisCenteringInclusivityDesign2021, sarabipourChangingScientificMeetings2021}.
Virtual conferences remove barriers for inclusion like costs of participation, the logistics of long-distance travel, and discriminatory visa applications \cite{jooKeepOnlineOption2021, ninerBetterWhomLeveling2021, salibaGettingGripsOnline2020, gichoraTenSimpleRules2010a}. 
Make sure that you are not creating additional barriers when choosing conference platforms, schedule, and activities. 
Check every platform's screen reader accessibility (see \textbf{Rule 6}), technological requirements (a high-speed internet or a latest generation computer prerequisite could exclude attendees from many regions), and geopolitical restrictions (some services are not available in all countries).
Be mindful of the different time zones and time constraints of the potential attendees and create flexible schedules.
Avoid creating a strictly synchronous live program, ask for pre-recorded talks, record sessions, and make them available during the conference.

@@

Conferences are among the least accessible spaces that people with disabilities may encounter in professional contexts \cite{priceAccessImaginedConstruction2009}. Even when conferences implement other inclusive practices, the participation of people with disabilities is often overlooked \cite{marks2021meeting}. Including people with disabilities in the organizing team from the start can have a large impact (see \textbf{Rule 3}), since planning for accessibility requires time, experience, and early decision-making, because inaccessible features are difficult to correct at later stages \cite{irishIncreasingParticipationUsing2020}. To ensure that disabled organizers can contribute substantially, the tools selected for behind-the-scenes organizing, communication, and planning the conference must be compatible with adaptive technology like screen readers, as well as providing captions for spoken conversations for deaf people. Many products currently available to organizers come with barriers to full participation. The people who use these technologies have expertise on which products work best with them.

If the conference has an in-person component, the venue should comply with common accessibility standards, such as being adequate for people who use wheelchairs, having signs in Braille, and using a sound system compatible with hearing devices and language interpretation, to name a few. 
Invisible disabilities  (e.g. dyslexia, anxiety, ADHD, autism) should be accounted for proactively, for example, by providing quiet spaces for privacy and noise-free conversations, or providing chairs in open spaces.

@@

Encourage the preparation of accessible slides and presentations by providing accessibility guidelines and presentation templates and be available for any questions that presenters and attendees may have; see \cite{sanchez-tapia_user_2021} for an example of accessibility guidelines and \cite{chavez_preparing_2021, sanchez-tapia_making_2021, joo_how_2021} for further accessibility recommendations.
Social events and networking should also have accessibility features like captioning or sign language interpretation, and include activities that do not restrict participation based on body type or ability.

@@

The communication strategy of your conference should aim to reach broader audiences and express the commitment with diversity and inclusion and the welcoming spirit of the conference.
Actively reach out and promote the conference to communities that have been systematically excluded. 
Promote the abstract submission call beyond the usual communication channels and reach out to local groups and communities of practice to encourage submissions by their members.
The members of your team should be able to decide which languages to emphasize and which social media platforms to utilize for the promotion of the conference (e.g. Twitter, Facebook, LinkedIn, conference website, mailing lists).

@@

Use inclusive language—language free from words, phrases, or tones that reflect prejudiced or discriminatory views of particular people or groups—in all communications \cite{hallDesigningDiversityInclusion2019}. 
Avoid as well all offensive or sexualized imagery in slides, conference materials, social events, and.
Become familiar with the terminology used for disabilities, racialized groups, gender and sexual orientations, terms that are preferred by each group and the terms that should be avoided.
Do not expect minoritized people to teach you and accept feedback without being offended.
Inclusive language also encompasses avoidance of excessive and too specific technical jargon and acronyms.

@@

% Rocío: main idea: help take the burden out of participants for more inclusive participation
When determining the registration rates, the socioeconomic context of participants, their country of origin, and their career status should be taken into account  \cite{sarabipourChangingScientificMeetings2021, andalibPostdocQueueLabour2018, kaplanPostdocNot2012}
(see \cite{canelon2021cost} for an example of conversion rates based on country of origin and career status). 
Resources permitting, enable a `pay what you can' option. You could also aim to have a conference with no registration costs; this is especially true for online events, but bear in mind that free events have a lower attendance rate than paid events \cite{eventbrite_ultimate_2017}. 
Scholarships to attend the in-person component of the conference are an additional way to boost participation of people from minoritized groups, by offering support for travel and lodging expenses.
Design the diversity-related scholarship to be explicit about the groups you want to support, and be transparent about the criteria for evaluation to avoid self-selection (refraining from applying). 
To assign these scholarships, conferences usually ask for cover letters or applications, which can be time-consuming and emotionally demanding. 
Simplifying the process of requesting financial support, especially for 
people who lack the time and resources (e.g. because of family responsibilities) could greatly increase their chances of applying for the support. 
Additional aid for attendees could also be offered—e.g. paying for child care or internet connection services for the online component. 
Bear in mind that transferring money internationally could be a cumbersome administrative process and significantly augment the burden of already marginalized groups. Whenever possible, facilitate alternate ways for money transfer (e.g. book flights, hotel reservations, waive conference fees, give gift cards). Depending on the country, services like OpenCollective (\url{https://opencollective.com}) can be very useful to do these transfers.

@@

Be aware of all the factors that may hinder your efforts for inclusion: systemic discrimination, time or money constraints, geopolitical or public health contexts, or personal issues affecting your organizing team. 
Try to account for these factors as best as you can to set a realistic yet bold vision and goals of diversity and inclusion for the conference.
At the end of the day, even if you do not achieve them at 100\%, striving for them will make you stretch what feels possible and help you make changes you would have otherwise not even attempted to make. 
And most importantly, you would have contributed to a great conference experience for the attendees. 
They will notice the welcoming and inclusive spirit of the conference. 
Inclusive conferences can have lasting impacts in career paths and create a sense of belonging in the community. 
That is more than enough motivation to make the effort worth it.

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 (\url{https://user2021.r-project.org/about/global-team}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process. Special thanks to Saranjeet Kaur and Gwynn Sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures.
